---
title: "Introduction to occumb"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to occumb}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  comment = "#>"
)
```

```{r, echo = FALSE, message = FALSE}
library(occumb)
set.seed(1)
```

This section briefly describes the steps involved in a series of model-based analyses of eDNA metabarcoding data that can be performed using the `occumb` R package.

## Prepare dataset

We will use the package's built-in data `fish` (see `?fish` for documentation) to see how to analyze sequence count data with the `occumb` package.
The `summary()` function can be used to display an overview of the dataset.

```{r, eval = FALSE}
data(fish)
summary(fish)
```

```{r}
#> Sequence read counts: 
#>  Number of species, I = 50 
#>  Number of sites, J = 50 
#>  Maximum number of replicates per site, K = 3 
#>  Number of missing observations = 6 
#>  Number of replicates per site: 2.88 (average), 0.33 (sd)
#>  Sequencing depth: 77910 (average), 98034.7 (sd)
#> 
#> Species covariates: 
#>  mismatch (continuous) 
#> Site covariates: 
#>  riverbank (categorical) 
#> Replicate covariates: 
#>  (None) 
#> 
#> Labels for species: 
#>  Abbottina rivularis, Acanthogobius lactipes, Acheilognathus macropterus, Acheilognathus rhombeus, Anguilla japonica, Biwia zezera, Carassius cuvieri, Carassius spp., Channa argus, Ctenopharyngodon idella, Cyprinus carpio, Gambusia affinis, Gnathopogon spp., Gymnogobius castaneus, Gymnogobius petschiliensis, Gymnogobius urotaenia, Hemibarbus spp., Hypomesus nipponensis, Hypophthalmichthys spp., Hyporhamphus intermedius, Ictalurus punctatus, Ischikauia steenackeri, Lepomis macrochirus macrochirus, Leucopsarion petersii, Megalobrama amblycephala, Micropterus dolomieu dolomieu, Micropterus salmoides, Misgurnus spp., Monopterus albus, Mugil cephalus cephalus, Mylopharyngodon piceus, Nipponocypris sieboldii, Nipponocypris temminckii, Opsariichthys platypus, Opsariichthys uncirostris uncirostris, Oryzias latipes, Plecoglossus altivelis altivelis, Pseudogobio spp., Pseudorasbora parva, Rhinogobius spp., Rhodeus ocellatus ocellatus, Salangichthys microdon, Sarcocheilichthys variegatus microoculus, Silurus asotus, Squalidus chankaensis biwae, Tachysurus tokiensis, Tanakia lanceolata, Tribolodon brandtii maruta, Tribolodon hakonensis, Tridentiger spp. 
#> Labels for sites: 
#>  1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13, 14, 15, 16, 17, 18, 19, 20, 21, 22, 23, 24, 25, 26, 27, 28, 29, 30, 31, 32, 33, 34, 35, 36, 37, 38, 39, 40, 41, 42, 43, 44, 45, 46, 47, 48, 49, 50 
#> Labels for replicates: 
#>  L, C, R 
```

Note that a summary of the sequence read count data is found in the first block of the output.
It shows that the data was obtained from samples of 50 sites * (up to) 3 replicates, with 50 fish species recorded.
A summary of the missing samples, number of replicates per site, and sequencing depth (i.e., the total number of sequence reads per sample) are also presented.
The second block of the output indicates that `fish` data also has two covariates, one continuous species covariate, `mismatch`, and one discrete site covariate, `riverbank`, which can be used in the following analysis.
The third block of the output shows labels given to the 50 species, 50 sites, and three replicates.

As such, the `fish` data brings together a set of data relevant to the analysis with the `occumb` package.
When you are analyzing your own dataset, you can use `occumbData()` to set up your data object.

If you want to display the entire data, just type `fish`.

```{r, eval = FALSE}
fish
```

## Fit models

With the `fish` dataset (or if your dataset is constructed with `occumbData()`), you can fit a multispecies site occupancy model for eDNA metabarcoding using the `occumb()` function.

The `occumb()` function has several arguments that accept model formulas to fit different model variants using the usual R formula syntax. See `vignette("model_specification")` for an overview of the models that `occumb()` can fit and details on how to specify models.

Fitting a null model (i.e., an intercept-only model, though the value of the intercept is species-specific) is very simple, as follows.

```{r, eval = FALSE}
fit0 <- occumb(data = fish, parallel = TRUE)
```

Setting `parallel = TRUE` is recommended for faster model fitting through parallel computation. A model incorporating species and site covariates in the `fish` dataset can be fitted, for example, as follows.

```{r, eval = FALSE}
fit1 <- occumb(formula_psi = ~ riverbank,
               formula_phi_shared = ~ mismatch,
               data = fish,
               parallel = TRUE)
```

Here, `riverbank` is specified as the covariate for site occupancy probability (`psi`) in the `formula_psi` argument and `mismatch` as the covariate for relative dominance of sequence (`phi`) in the `formula_phi_shared` argument. In the latter, since `mismatch` is a species covariate, the `formula_phi_shared` argument is used instead of `formula_phi` so that the effect of `mismatch` is constant across species; see `vignette("model_specification")` for more details.

The above two model fits can take approximately one hour. Given that `occumb()` fits a class of complex hierarchical models with many parameters and unknown variables, it may require, depending on the dataset size, a long time and/or a lot of memory for model fitting.

`occumb()` also has several arguments to control the Markov Chain Monte Carlo (MCMC) computation for fitting the model. For example, to obtain more precise posterior estimates than the example above, you can explicitly set the `n.iter` and `n.thin` arguments to get longer and less autocorrelated MCMC samples.

```{r, eval = FALSE}
fit1x <- occumb(formula_psi = ~ riverbank,
                formula_phi_shared = ~ mismatch,
                data = fish,
                n.thin = 20,
                n.iter = 40000,
                parallel = TRUE)
```


## Check posterior samples

With the `fit1` object, we will look at how to access and diagnose the posterior sample obtained using `occumb()`.

The `print()` function summarizes the MCMC results, including a table of posterior estimates for each saved parameter. The `plot()` function draws a trace plot and a density plot for each parameter. Although results are not shown here, users who have used the `{jagsUI}` package should be familiar with the output of these functions. Model fitting in `occumb()` relies on functions provided by the `{jagsUI}` package, and the `plot()` and `print()` functions apply the methods for `{jagsUI}` objects.

```{r, eval = FALSE}
print(fit1)
plot(fit1)
```

The `summary()` function provides a high-level summary of the result. This is based on the summary method for `{jagsUI}` objects and modified for the `occumb()` output, which displays data summary, model equations, MCMC settings, and a quick overview of the convergence of relevant model parameters.

```{r, eval = FALSE}
summary(fit1)
```

```{r}
#> Summary for occumbFit object 
#> 
#> Summary of data:
#>  Number of species, I = 50 
#>  Number of sites, J = 50 
#>  Maximum number of replicates per site, K = 3 
#>  Number of missing observations = 6 
#>  Number of replicates per site: 2.88 (average), 0.33 (sd) 
#>  Sequencing depth: 77910 (average), 98034.7 (sd) 
#> 
#> Model specification:
#>  formula_phi:          ~ 1 
#>  formula_theta:        ~ 1 
#>  formula_psi:          ~ riverbank 
#>  formula_phi_shared:   ~ mismatch 
#>  formula_theta_shared: ~ 1 
#>  formula_psi_shared:   ~ 1 
#>  prior_prec:           1e-04 
#>  prior_ulim:           10000 
#> 
#> Saved parameters:
#>  Mu sigma rho alpha beta gamma alpha_shared phi theta psi z pi deviance 
#> 
#> MCMC ran in parallel for 46.39 minutes at time 2023-07-13 15:44:34.56973:
#>  For each of 4 chains:
#>   Adaptation:            100 iterations (sufficient)
#>   Burn-in:               10000 iterations
#>   Thin rate:             10 iterations
#>   Total chain length:    20100 iterations
#>   Posterior sample size: 1000 draws
#> 
#> Summary of posterior samples: 
#>  Mu: 
#>   Number of parameters: 4 
#>   Rhat:  1.001 (min), 1.008 (median), 1.046 (mean), 1.168 (max) 
#>   n.eff: 20 (min), 378 (median), 814.2 (mean), 2481 (max) 
#>  sigma: 
#>   Number of parameters: 4 
#>   Rhat:  1.011 (min), 1.046 (median), 1.046 (mean), 1.081 (max) 
#>   n.eff: 37 (min), 141.5 (median), 167 (mean), 348 (max) 
#>  rho: 
#>   Number of parameters: 16 
#>   Rhat:  1 (min), 1.001 (median), 1.028 (mean), 1.222 (max) 
#>   n.eff: 17 (min), 1772 (median), 2039.5 (mean), 4000 (max) 
#>  alpha: 
#>   Number of parameters: 50 
#>   Rhat:  1.008 (min), 1.158 (median), 1.205 (mean), 2.248 (max) 
#>   n.eff: 6 (min), 23.5 (median), 48.6 (mean), 466 (max) 
#>  beta: 
#>   Number of parameters: 50 
#>   Rhat:  1.002 (min), 1.02 (median), 1.029 (mean), 1.117 (max) 
#>   n.eff: 36 (min), 160 (median), 269.2 (mean), 1768 (max) 
#>  gamma: 
#>   Number of parameters: 100 
#>   Rhat:  1 (min), 1.005 (median), 1.007 (mean), 1.062 (max) 
#>   n.eff: 51 (min), 599.5 (median), 950 (mean), 4000 (max) 
#>  alpha_shared: 
#>   Number of parameters: 1 
#>   Rhat:  1.402 
#>   n.eff: 11 
#>  phi: 
#>   Number of parameters: 50 
#>   Rhat:  1.002 (min), 1.053 (median), 1.158 (mean), 2.5 (max) 
#>   n.eff: 5 (min), 54 (median), 167.1 (mean), 4000 (max) 
#>  theta: 
#>   Number of parameters: 50 
#>   Rhat:  1.002 (min), 1.018 (median), 1.024 (mean), 1.084 (max) 
#>   n.eff: 39 (min), 173 (median), 273.2 (mean), 1820 (max) 
#>  psi: 
#>   Number of parameters: 2500 
#>   Rhat:  1 (min), 1.003 (median), 1.006 (mean), 1.047 (max) 
#>   n.eff: 65 (min), 894 (median), 1401.7 (mean), 4000 (max) 
#>  z: 
#>   Number of parameters: 2500 
#>   Rhat:  1 (min), 1.005 (median), 1.009 (mean), 1.099 (max), 946 (Number of NAs) 
#>   n.eff: 1 (min), 1220 (median), 1665.4 (mean), 4000 (max) 
#>  pi: 
#>   Number of parameters: 7500 
#>   Rhat:  1 (min), 1.077 (median), 1.1 (mean), 1.293 (max), 28 (Number of NAs) 
#>   n.eff: 1 (min), 1 (median), 1016.5 (mean), 4000 (max) 
#>  deviance: 
#>   Rhat:  1.001 
#>   n.eff: 3259 
#> 
#> **WARNING** Rhat values indicate convergence failure (Rhat > 1.1). 
#> 
#> DIC info: (pD = var(deviance)/2)
#>  pD = 2152.9 and DIC = 19993.73 
```

Overall, the results are close to convergence, but the convergence of `phi` and `alpha`, the parameters related to the relative dominance of species sequence reads, is not good (i.e., high values of `Rhat`); hence, longer MCMC runs will be required for a formal inference.

If you want to access posterior samples of individual parameters or their summaries, you can use the `get_post_samples()` and `get_post_summary()` functions. As an example, try to get a posterior summary of site occupancy probability `psi` using the `get_post_summary()` function.

```{r, eval = FALSE}
post_summary_psi <- get_post_summary(fit1, "psi")
post_summary_psi
```

```{r}
#>                  mean          sd        2.5%        25%        50%        75%      97.5%     Rhat n.eff overlap0 f
#> psi[1,1]   0.64607375 0.083034133 0.466158634 0.59382744 0.65078275 0.70478301 0.79318634 1.001697  1264        0 1
#> psi[2,1]   0.28250539 0.089275014 0.136970706 0.21789079 0.27480943 0.33672122 0.47971573 1.013796   317        0 1
#> psi[3,1]   0.38321762 0.088583160 0.218908875 0.32150169 0.38055944 0.44171514 0.56510758 1.005127   472        0 1
#> psi[4,1]   0.16899696 0.070833671 0.058363836 0.11684541 0.15943020 0.21075828 0.32821669 1.001143  3788        0 1
#> psi[5,1]   0.20435372 0.088327893 0.069729103 0.13950376 0.19315615 0.25286705 0.41681236 1.006308   596        0 1
#> psi[6,1]   0.10737782 0.061179251 0.024299550 0.06246265 0.09573034 0.13925811 0.25577838 1.002291  1418        0 1
#> psi[7,1]   0.84639119 0.053478041 0.730770215 0.81327035 0.85048238 0.88494478 0.93751979 1.003575   654        0 1
#> psi[8,1]   0.94296846 0.028288919 0.875373936 0.92739641 0.94770976 0.96354354 0.98337253 1.002581  1006        0 1
#> (The rest omitted)
```

In the `fit1` model, `psi` is assumed to have different values for each site depending on the `riverbank` covariate. Therefore, `psi` is estimated for each 50 species $\times$ 50 sites (note that `psi` differs for each species by default), and two subscripts in parentheses distinguish them. So, which species and site parameter is `psi[1,1]`? How about `psi[2,1]`? Information about the dimension of the parameter is given in the attributes of the resulting object.

```{r, eval = FALSE}
attributes(post_summary_psi)$dimension
```

```{r}
#> [1] "Species" "Site"
```

```{r, eval = FALSE}
attributes(post_summary_psi)$label
```

```{r}
#> $Species
#>  [1] "Abbottina rivularis"                     
#>  [2] "Acanthogobius lactipes"                  
#>  [3] "Acheilognathus macropterus"              
#>  [4] "Acheilognathus rhombeus"                 
#>  [5] "Anguilla japonica"                       
#>  [6] "Biwia zezera"                            
#>  [7] "Carassius cuvieri"                       
#>  [8] "Carassius spp."                          
#>  [9] "Channa argus"                            
#> [10] "Ctenopharyngodon idella"                 
#> [11] "Cyprinus carpio"                         
#> [12] "Gambusia affinis"                        
#> [13] "Gnathopogon spp."                        
#> [14] "Gymnogobius castaneus"                   
#> [15] "Gymnogobius petschiliensis"              
#> [16] "Gymnogobius urotaenia"                   
#> [17] "Hemibarbus spp."                         
#> [18] "Hypomesus nipponensis"                   
#> [19] "Hypophthalmichthys spp."                 
#> [20] "Hyporhamphus intermedius"                
#> [21] "Ictalurus punctatus"                     
#> [22] "Ischikauia steenackeri"                  
#> [23] "Lepomis macrochirus macrochirus"         
#> [24] "Leucopsarion petersii"                   
#> [25] "Megalobrama amblycephala"                
#> [26] "Micropterus dolomieu dolomieu"           
#> [27] "Micropterus salmoides"                   
#> [28] "Misgurnus spp."                          
#> [29] "Monopterus albus"                        
#> [30] "Mugil cephalus cephalus"                 
#> [31] "Mylopharyngodon piceus"                  
#> [32] "Nipponocypris sieboldii"                 
#> [33] "Nipponocypris temminckii"                
#> [34] "Opsariichthys platypus"                  
#> [35] "Opsariichthys uncirostris uncirostris"   
#> [36] "Oryzias latipes"                         
#> [37] "Plecoglossus altivelis altivelis"        
#> [38] "Pseudogobio spp."                        
#> [39] "Pseudorasbora parva"                     
#> [40] "Rhinogobius spp."                        
#> [41] "Rhodeus ocellatus ocellatus"             
#> [42] "Salangichthys microdon"                  
#> [43] "Sarcocheilichthys variegatus microoculus"
#> [44] "Silurus asotus"                          
#> [45] "Squalidus chankaensis biwae"             
#> [46] "Tachysurus tokiensis"                    
#> [47] "Tanakia lanceolata"                      
#> [48] "Tribolodon brandtii maruta"              
#> [49] "Tribolodon hakonensis"                   
#> [50] "Tridentiger spp."                        
#> 
#> $Site
#>  [1] "1"  "2"  "3"  "4"  "5"  "6"  "7"  "8"  "9"  "10" "11" "12" "13" "14" "15"
#> [16] "16" "17" "18" "19" "20" "21" "22" "23" "24" "25" "26" "27" "28" "29" "30"
#> [31] "31" "32" "33" "34" "35" "36" "37" "38" "39" "40" "41" "42" "43" "44" "45"
#> [46] "46" "47" "48" "49" "50"
```

Thus, from the `$dimension` attribute, we know that the first subscript of `psi` represents the species and the second the site. From the `$label` attribute, we know that `psi[1,1]` represents the value of `psi` for *Abbottina rivularis* at site 1, and `psi[2,1]` represents the value of `psi` for *Acanthogobius lactipes* at site 1.

We will also look at the posterior summary of the effect term `gamma` on `psi`.

```{r, eval = FALSE}
post_summary_gamma <- get_post_summary(fit1, "gamma")
attributes(post_summary_gamma)$dimension
```

```{r}
#> [1] "Species" "Effects"
```

```{r, eval = FALSE}
attributes(post_summary_gamma)$label$Effects
```

```{r}
#> [1] "(Intercept)"                 "riverbankwithout_vegetation"
```

`gamma` has two effect values for each species: the intercept `(Intercept)` and the effect of the absence of vegetation in the riverbank `riverbankwithout_vegetation`. Displaying the summary (results not shown below), we will see that, for many species, the absence of vegetation has a negative effect on `psi` (on its link scale).

```{r, eval = FALSE}
post_summary_gamma
```

Next, we will access posterior samples of `psi` using the `get_post_samples()` function.

```{r, eval = FALSE}
post_sample_psi <- get_post_samples(fit1, "psi")
dim(post_sample_psi)
```

```{r}
#> [1] 4000   50   50
```

The resulting object `post_sample_psi` is a 3-dimensional array containing the posterior sample of `psi`. Information about the dimensions of the extracted sample is given in its attributes.

```{r, eval = FALSE}
attributes(post_sample_psi)
```

```{r}
#> $dim
#> [1] 4000   50   50
#> 
#> $dimension
#> [1] "Sample"  "Species" "Site"   
#> 
#> $label
#> $label$Sample
#> NULL
#> 
#> $label$Species
#>  [1] "Abbottina rivularis"                     
#>  [2] "Acanthogobius lactipes"                  
#>  [3] "Acheilognathus macropterus"              
#>  [4] "Acheilognathus rhombeus"                 
#>  [5] "Anguilla japonica"                       
#>  [6] "Biwia zezera"                            
#>  [7] "Carassius cuvieri"                       
#>  [8] "Carassius spp."                          
#>  [9] "Channa argus"                            
#> [10] "Ctenopharyngodon idella"                 
#> [11] "Cyprinus carpio"                         
#> [12] "Gambusia affinis"                        
#> [13] "Gnathopogon spp."                        
#> [14] "Gymnogobius castaneus"                   
#> [15] "Gymnogobius petschiliensis"              
#> [16] "Gymnogobius urotaenia"                   
#> [17] "Hemibarbus spp."                         
#> [18] "Hypomesus nipponensis"                   
#> [19] "Hypophthalmichthys spp."                 
#> [20] "Hyporhamphus intermedius"                
#> [21] "Ictalurus punctatus"                     
#> [22] "Ischikauia steenackeri"                  
#> [23] "Lepomis macrochirus macrochirus"         
#> [24] "Leucopsarion petersii"                   
#> [25] "Megalobrama amblycephala"                
#> [26] "Micropterus dolomieu dolomieu"           
#> [27] "Micropterus salmoides"                   
#> [28] "Misgurnus spp."                          
#> [29] "Monopterus albus"                        
#> [30] "Mugil cephalus cephalus"                 
#> [31] "Mylopharyngodon piceus"                  
#> [32] "Nipponocypris sieboldii"                 
#> [33] "Nipponocypris temminckii"                
#> [34] "Opsariichthys platypus"                  
#> [35] "Opsariichthys uncirostris uncirostris"   
#> [36] "Oryzias latipes"                         
#> [37] "Plecoglossus altivelis altivelis"        
#> [38] "Pseudogobio spp."                        
#> [39] "Pseudorasbora parva"                     
#> [40] "Rhinogobius spp."                        
#> [41] "Rhodeus ocellatus ocellatus"             
#> [42] "Salangichthys microdon"                  
#> [43] "Sarcocheilichthys variegatus microoculus"
#> [44] "Silurus asotus"                          
#> [45] "Squalidus chankaensis biwae"             
#> [46] "Tachysurus tokiensis"                    
#> [47] "Tanakia lanceolata"                      
#> [48] "Tribolodon brandtii maruta"              
#> [49] "Tribolodon hakonensis"                   
#> [50] "Tridentiger spp."                        
#> 
#> $label$Site
#>  [1] "1"  "2"  "3"  "4"  "5"  "6"  "7"  "8"  "9"  "10" "11" "12" "13" "14" "15"
#> [16] "16" "17" "18" "19" "20" "21" "22" "23" "24" "25" "26" "27" "28" "29" "30"
#> [31] "31" "32" "33" "34" "35" "36" "37" "38" "39" "40" "41" "42" "43" "44" "45"
#> [46] "46" "47" "48" "49" "50"
```

Thus, the second dimension of `post_sample_psi` corresponds to the species and the third to the site. The labels for species and sites indicate that `post_sample_psi[, 1, 1]` contains the posterior sample of `psi` of *Abbottina rivularis* at site 1.

```{r, eval = FALSE}
post_sample_psi[, 1, 1]
```

```{r}
#>  [1] 0.6673664 0.7505735 0.6416083 0.7011978 0.5935338 0.6087940 0.6159829
#>  [8] 0.7758206 0.6710466 0.7423708 0.6244308 0.6237291 0.6344345 0.6715390
#> [15] 0.6255393 0.7306071 0.6781538 0.7391794 0.6678208 0.6715868 0.7582302
#> [22] 0.5947020 0.4880676 0.6958859 0.6202520 0.5487584 0.6781225 0.7073072
#> (The rest omitted)
```



## Assess goodness-of-fit

Assessing the goodness-of-fit of the fitted model is an essential step in data analysis based on statistical models. The goodness-of-fit of models fitted with `occumb()` can be evaluated using the `gof()` function which computes the Bayesian p-value via the posterior predictive check approach.

```{r, eval = FALSE}
gof_result <- gof(fit1, cores = 4)
```

It is recommended to specify the `cores` argument explicitly for parallel computation.


## Analyze study design

The model fitted with `occumb()` can be used to find eDNA metabarcoding study designs that effectively detect species present. How many sites, within-site replicates, and sequencing depth will be required to ensure reliable species detection? What is the best balance between the number of sites visited, the number of within-site replicates, and sequencing depth under a limited budget? The multispecies site occupancy model fit by `occumb()` answers these questions by predicting the number of species expected to be detected under specific study designs.

`eval_util_L()` and `eval_util_R()` are functions available for these purposes. These two functions assume species diversity assessment at different spatial scales (i.e., the former is for local (L), the latter for regional (R)). Specifically, `eval_util_L()` is appropriate if you are interested in assessing species diversity only for the study sites included in the dataset, and `eval_util_R()` is appropriate if you are interested in assessing species diversity in a broader area that includes the study sites in the dataset (i.e., the population of sites or a "metacommunity").

In the context of the `fish` dataset, `eval_util_L()` evaluates the expected number of species detected per site at the 50 sites in the dataset under different combinations of the number of replicates and sequencing depth. For example, the following will give the expected number of species detected per site when the number of replicates ($K$) and the sequencing depth ($N$) take the values (1, 2, 3) and (1000, 10000, 100000), respectively.

```{r, eval = FALSE}
utilL1 <- eval_util_L(expand.grid(K = 1:3, N = c(1E3, 1E4, 1E5)),
                      fit1, cores = 4)
utilL1
```

```{r}
#>   K     N  Utility
#> 1 1 1e+03 13.03225
#> 2 2 1e+03 16.72177
#> 3 3 1e+03 18.21783
#> 4 1 1e+04 14.34826
#> 5 2 1e+04 17.72304
#> 6 3 1e+04 18.92009
#> 7 1 1e+05 15.04591
#> 8 2 1e+05 18.20935
#> 9 3 1e+05 19.23478
```

Note that it is recommended to specify the `cores` argument explicitly for parallel computation.

If the cost per sequence read for high-throughput sequencing, the cost per replicate for library preparation, and the research budget values are each known, then the `list_cond_L()` function can be used to obtain a set of feasible settings under these cost and budget values. This can be given to `eval_util_L()` to identify the optimal study design under a budget constraint.

```{r, eval = FALSE}
settings <- list_cond_L(budget = 875 * 1E3,
                        lambda1 = 0.01,
                        lambda2 = 5000,
                        fit1)
utilL2 <- eval_util_L(settings, fit1, cores = 4)
utilL2
```

```{r}
#>   budget lambda1 lambda2 K          N  Utility
#> 1 875000    0.01    5000 1 1250000.00 15.45671
#> 2 875000    0.01    5000 2  375000.00 18.37862
#> 3 875000    0.01    5000 3   83333.33 19.21953
```

Hence, under this specific budget and cost value, it is best to have K = 3 replicates per site for the most effective species detection.

In contrast to `eval_util_L()`, `eval_util_R()` considers differences in the number of sites visited. Nevertheless, `eval_util_R()` can be applied in much the same way as `eval_util_L()`. Since `list_cond_R()` may return a huge list of possible settings, you may want to get a manually restricted list using its `J` and/or `K` arguments.

```{r, eval = FALSE}
# Use fit0 here since the current version of eval_util_R() does not support
# site heterogeneity of occupancy probability.
utilR1 <- eval_util_R(expand.grid(J = 1:3, K = 1:3, N = c(1E3, 1E4, 1E5)),
                      fit0, cores = 4)
utilR1
```

```{r}
#>    J K     N  Utility
#> 1  1 1 1e+03 13.02089
#> 2  2 1 1e+03 18.75257
#> 3  3 1 1e+03 22.28694
#> 4  1 2 1e+03 16.75038
#> 5  2 2 1e+03 22.84567
#> 6  3 2 1e+03 26.37878
#> 7  1 3 1e+03 18.22010
#> 8  2 3 1e+03 24.50571
#> 9  3 3 1e+03 28.16144
#> 10 1 1 1e+04 14.39106
#> 11 2 1 1e+04 20.38495
#> 12 3 1 1e+04 23.88514
#> 13 1 2 1e+04 17.72739
#> 14 2 2 1e+04 24.01112
#> 15 3 2 1e+04 27.58758
#> 16 1 3 1e+04 18.95701
#> 17 2 3 1e+04 25.38315
#> 18 3 3 1e+04 29.11239
#> 19 1 1 1e+05 14.99955
#> 20 2 1 1e+05 21.19485
#> 21 3 1 1e+05 24.77456
#> 22 1 2 1e+05 18.17301
#> 23 2 2 1e+05 24.54592
#> 24 3 2 1e+05 28.28593
#> 25 1 3 1e+05 19.18010
#> 26 2 3 1e+05 25.78404
#> 27 3 3 1e+05 29.49870
```

```{r, eval = FALSE}
settings <- list_cond_R(budget = 1125 * 1E3,
                        lambda1 = 0.01,
                        lambda2 = 5000,
                        lambda3 = 5000,
                        J = seq(5, 50, 5),
                        K = 1:4)
utilR2 <- eval_util_R(settings, fit0, cores = 4)
utilR2
```

```{r}
#>     budget lambda1 lambda2 lambda3  J K           N  Utility
#> 1  1125000    0.01    5000    5000  5 1 21500000.00 30.10762
#> 2  1125000    0.01    5000    5000 10 1 10250000.00 35.66881
#> 3  1125000    0.01    5000    5000 15 1  6500000.00 38.58941
#> 4  1125000    0.01    5000    5000 20 1  4625000.00 40.54363
#> 5  1125000    0.01    5000    5000 25 1  3500000.00 41.97652
#> 6  1125000    0.01    5000    5000 30 1  2750000.00 43.03573
#> 7  1125000    0.01    5000    5000 35 1  2214285.71 43.86356
#> 8  1125000    0.01    5000    5000 40 1  1812500.00 44.63576
#> 9  1125000    0.01    5000    5000 45 1  1500000.00 45.15320
#> 10 1125000    0.01    5000    5000 50 1  1250000.00 45.60239
#> 11 1125000    0.01    5000    5000  5 2 10500000.00 33.26731
#> 12 1125000    0.01    5000    5000 10 2  4875000.00 38.67689
#> 13 1125000    0.01    5000    5000 15 2  3000000.00 41.48350
#> 14 1125000    0.01    5000    5000 20 2  2062500.00 43.33658
#> 15 1125000    0.01    5000    5000 25 2  1500000.00 44.57691
#> 16 1125000    0.01    5000    5000 30 2  1125000.00 45.48365
#> 17 1125000    0.01    5000    5000 35 2   857142.86 46.19298
#> 18 1125000    0.01    5000    5000 40 2   656250.00 46.72431
#> 19 1125000    0.01    5000    5000 45 2   500000.00 47.16480
#> 20 1125000    0.01    5000    5000 50 2   375000.00 47.46552
#> 21 1125000    0.01    5000    5000  5 3  6833333.33 34.49460
#> 22 1125000    0.01    5000    5000 10 3  3083333.33 39.92865
#> 23 1125000    0.01    5000    5000 15 3  1833333.33 42.70624
#> 24 1125000    0.01    5000    5000 20 3  1208333.33 44.45794
#> 25 1125000    0.01    5000    5000 25 3   833333.33 45.64077
#> 26 1125000    0.01    5000    5000 30 3   583333.33 46.42328
#> 27 1125000    0.01    5000    5000 35 3   404761.90 47.04303
#> 28 1125000    0.01    5000    5000 40 3   270833.33 47.47639
#> 29 1125000    0.01    5000    5000 45 3   166666.67 47.80421
#> 30 1125000    0.01    5000    5000 50 3    83333.33 48.02496
#> 31 1125000    0.01    5000    5000  5 4  5000000.00 35.00271
#> 32 1125000    0.01    5000    5000 10 4  2187500.00 40.55395
#> 33 1125000    0.01    5000    5000 15 4  1250000.00 43.37786
#> 34 1125000    0.01    5000    5000 20 4   781250.00 45.00144
#> 35 1125000    0.01    5000    5000 25 4   500000.00 46.10021
#> 36 1125000    0.01    5000    5000 30 4   312500.00 46.86324
#> 37 1125000    0.01    5000    5000 35 4   178571.43 47.40744
#> 38 1125000    0.01    5000    5000 40 4    78125.00 47.80381
```

